Speculation in foreign exchange: noise or information

### 2015-09-15 Knitr/NewPaper.Rnw
Updated the NewPaper and Extreme2.R The paper is ready for Andros, the R code will loop thorugh the extremes and create charts of the distribution post extreme. There is also a comparison of extreme move and the level of speculative sentiment.  Table produced.

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
 Uninformed speculation based on noise should lead to price reversals; informed speculation is sustainable. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal.  Price activity after the extreme is close to a random walk. This supports the view that speculation is part of the price-discovery process.      
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial market span a broad range.  Towards one end is the Keynesian view that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment; It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.

Towards the other end of the spectrum, Friedman argued that speculation was a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

@@

This paper contributes to the debate about the nature of speculation in a number of ways. A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.  Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk; this suggests that speculation is part of the process of price discovery.

For example, Figure \ref{fig:ES1} show the evolution of a selection of foreign exchange prices after an extreme reading of speculative sentiment. Sentiment is measured by an extreme in the price of currency call options relative to the equivalent put.  It is clear that the foreign exchange rate has a contemporaneous movement with the sentiment but that after reading the extreme, there is no clear direction.  The lack of clear direction indicates lack of information.  As such, extreme speculative positions do presage reversals, indicating that they are not associated with divergence from equilibrium.

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.6]{RRCum16}
\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns.  The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES1}
\end{figure}
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation is central to many contemporary questions about international financial markets. For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.  If speculation is largely uninformed it will add to the level of noise and maker it even harder for informed-traders to overcome uncertainty and apply fundamental information about long-term value. It will increase the tendency towards activity that is focused on the short-term.  However, if speculation is informed and part of the process of price discovery, regulatory measures may reduce liquidity and undermine the prices of price discovery. Increased volatility, heightened risk and reduced transparency together with other costs may provide an countervailing force.

The rest of the paper proceeds as follows:  Section Two analyses the distinction between informed and uninformed speculators; Section Three discusses the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.

\section{Informed and uninformed traders}
Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities and rational expectations. However, there is no clear, unambiguous theory of speculation.  While speculation may be vowed as being informed or uninformed, those trading without information are often categorised as being either \emph{noise-traders} or \emph{liquidity traders}.

Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.  Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Market-participants do not make decisions independently of those made by others. Information is inferred from the activities of others. \citet[p.101]{Keynes1936} said investment decisions could be based on the expectations of others' valuations.  institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''.

An important part of microstructure research is the distinction between informed and uninformed traders.  The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.  This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.  See also \citet{glosten1985bid}.

As such, Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to trade. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  Indeed, this space can encourage informed traders to pay the cost of acquiring information.  See, for example \citet{Grossman1980Impossibility}.

@@

\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.  The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; it may also be a function of the market-maker knowledge that large orders are more likely to informed traders seeking to take advantage of their informational advantage.  \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.  The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that this is more than offset by the high quantity of uninformed traders.  Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.

Informed speculation is also used as the theoretical justification for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.  Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate or sell forecasts that will tend to drive prices towards the equilibrium \citet{muth1961Rational}.  This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable. While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.

However, a large proportion of uninformed traders in the market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.  The risk that noise-traders will drive prices even further away from equilibrium before the convergence can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation} 
 where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho$ is the price mis-perception of the uninformed traders; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}

@@

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish between the informed and uninformed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  Common strategies that have been used are momentum (ref), contrarian (ref) and carry (ref).   It is also possible to assume that certain investors usually have more information than others.  For example, hedge funds, investment banks and other professionals can be assumed to be informed while retail or day traders may be uninformed.

%It has also been suggested that speculative techniques can change over time.  For example, \citet{ShillerFashion} discusses the use of fashionable techniques or fads in investment.  In this way, speculation can be be associated with sentiment and measures of sentiment are then used as ways of measuring speculation.

%Noise stragegies may be successful for a while.  Momentum trades tend to be successful for certain ties.  Also discssion the literature on heterogeneous traders, another way that the trading securities is explained, has very frequently used noise-traders as trend-chasers or followers of technical analysis.  These models (see for example brock and hommes [1998]; lux [1995]; lux and marchesi [2000]) investigate bubbles and crashes, and feature fundamental traders versus noise traders. 
\citet{EasleyPIN2} propose a model to try to measure informed market activity based on their assertion that the bid-ask spread is a function of the level of informed trading.  As informed traders are most likely to want to take advantage of their informational advantage, large orders are assumed to be informed.\footnote{With market-makers adopting Bayesian reasoning to update their fundamental valuation by using order flow to estimate the probability that there has been an information event and if there has whether it is good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

However, identifying large orders and making an assessment of the level of informed trading based on that is much harder to achieve in an over-the-counter market like foreign exchange. While electronic order books have become more widely used in recent years, much of the activity still takes place through bilateral agreements.  Information about the nature of transactions including size and origin of orders is not widely available\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.  They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent. They argue that the orders crystallise a range of information that is used to value currencies.}    \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.  Their tick test has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick. A trade is an up-tick if the price is higher than the previous trade.  If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick.  
\section{Methodology}
The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed. Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.  If informed and part of the price-discovery process, extreme speculation would be a function of perceived divergence from equilibrium or new information.  The following price performance should be a random walk.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.

Returning to the \citet[pp. 14-15]{Delong1990noise} model summarised by Equation \ref{eq:NP2}, there are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If either of these reach an extreme, prices should be expected to reverse\footnote{See \citet{Hayward2013} for full details of simulations of this model.}

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how strongly speculators believe that the market will move in a particular direction. In the absence of a daily survey of speculators, this study uses \emph{risk-reversal skew} as an objective measure of speculative sentiment rather than surveys of opinion.%\footnote{There are a number of surveys of investor opinion that are regularly conducted by financial service information companies.  However, these are disparate and, in many cases, unsystematic, and it is not clear whether they measure speculative sentiment or investor opinion more generally.   The Wall Street Journal runs an annual survey of investor opinion, but this is focused mostly on US equities.  The \emph{Economist} magazine also solicits opinion, but in a relatively unstructured and \emph{ad hoc} fashion.  In the foreign exchange market, there is a broad set of market forecasts that was collected by the financial services company \emph{MMS International} (formerly Money Market Services) which polled a broad range of financial market participants on key exchange rates for one and four week forecast horizons. Frankel and Froot found that the median forecast from the MMS survey exhibited trend-following behaviour at the one to three week horizon while there is mean-reversion beyond that point.  Frankel and Froot also assess the 'Financial Report' that has been produced by \emph{The Economist} magazine and \emph{Currency Forecasters Digest} and find a wide dispersion of opinion and express some concern about the aggregation of such divergent forecasters \citet gc{FrootFrankelFDB} and \citep{FrankelFroot1987}.}

Risk-reversal skew is the relative price of currency put and call options for the same strike price.  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters}.

This study uses a unique dataset of \emph{risk-reversals} for eleven major currency pairs.  It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

>>>>

\documentclass[12pt, a4paper, oneside]{article} % Paper size, default font size and one-sided paper
%\graphicspath{{./Figures/}} % Specifies the directory where pictures are stored
%\usepackage[dcucite]{harvard}
\usepackage{rotating}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{pdflscape}
\usepackage{appendix}
\usepackage[flushleft]{threeparttable}
\usepackage{multirow}
\usepackage[comma, sort&compress]{natbib}% Use the natbib reference package - read up on this to edit the reference style; if you want text (e.g. Smith et al., 2012) for the in-text references (instead of numbers), remove 'numbers' 
\usepackage{graphicx}
%\bibliographystyle{plainnat}
\bibliographystyle{agsm}
\usepackage[colorlinks = true, citecolor = blue, linkcolor = blue]{hyperref}
%\hypersetup{urlcolor=blue, colorlinks=true} % Colors hyperlinks in blue - change to black if annoying
%\renewcommand[\harvardurl]{URL: \url}
 \usepackage{listings}
  \usepackage{tikz}
 \usetikzlibrary{arrows,positioning}
 \usepackage{color}
 %\graphicspath{{../Pictures/}}
\definecolor{mygrey}{gray}{0.95}
\lstset{backgroundcolor=\color{mygrey}}
\begin{document}
\title{Speculation in foreign exchange: noise or information}
\author{Rob Hayward\footnote{University of Brighton Business School, Lewes Road, Brighton, BN2 4AT; Telephone 01273 642586.  rh49@brighton.ac.uk. }} 
\date{\today}
\maketitle
\begin{abstract}
If prices are driven by speculation based on noise rather than information, extreme levels of speculative sentiment and speculative activity should lead to price reversals. Informed speculation is sustainable as it is part of the price discovery process. A unique set of option prices measure speculative sentiment while regulatory positions determine the weight of speculative activity.  An event study shows that extremes of speculative sentiment or speculative activity do not increase the probability of price reversal, though variance-measured risk increases.  Price activity after the extreme is close to a random walk. This supports the view that speculation is part of the price-discovery process rather than noise.      
\end{abstract}
\section*{Introduction}
Ideas about the way that speculation affects financial market span a broad range.  Towards one end is the view of \citet[p. 101]{Keynes1936} that speculation is a myopic, sentiment-driven activity, dominated by the desire to ``beat the gun'' or  ``outwit the crowd'' \citet[p. 101]{Keynes1936}. Speculation is a problem that will cause booms, reversals and will discourage fundamental, long-term investment. It is noise that emerges from the interaction of collective sentiment to encourages trends in asset prices.

Towards the other end of the spectrum is the idea that speculation is a stabilising force, providing liquidity and helping to ensure that markets swiftly find equilibrium. \citet{FriedmanPositive} asserted that successful speculators would be those who managed to correctly buy when prices  were below fundamental value and sell above that price. Therefore, Friedman argued, the speculative process would reward those able to access or utilise information relative to the uninformed and, through a Darwinian process, informed speculation would dominate.  This view regards speculation as an key lubricant of financial markets as it provides liquidity and facilitates the process of price discovery.

@@

This paper contributes to the debate about the nature of speculation in a number of ways. A unique set of option prices are used to identify the intensity of speculative sentiment; an event study method is used to establish the effect of extremes of speculative sentiment and speculative positions; these speculative extremes are compared to a common noise-trader model.  Speculative extremes do not provide information about the future; the evolution of foreign exchange prices after an extreme can be characterised as a random walk. However, extremes are followed by an increase in the variance of exchange rate returns. This suggests that information from informed speculation contributes to the process of price discovery.

Figure \ref{fig:ES1} shows the evolution of a selection of foreign exchange prices after speculative sentiment is extreme. An event study is conducted with the speculative extreme being the event $(t_0)$,  Sentiment is measured by the price of currency call options relative to the equivalent put.  The foreign exchange rate has a contemporaneous movement with the sentiment.  This suggests that either the price is adjusting to new information or that uninformed speculators are driving the price in the direction of their irrational belief. However, when sentiment reaches an extreme, there is no clear direction: prices are as likely to go up as they are to go down.  As such, extreme speculative positions do not presage reversals, indicating that they are not associated with divergence from equilibrium.

\begin{figure}
\graphicspath{{../Figures/}}
\centering
\includegraphics[scale=0.8]{RRCum16}
\caption{Event Study:  Extreme (High 99th percentile) Risk Reversal Skew and 16 day event window: Risk reversals are the ratio of implied volatility on calls to the equivalent put; extreme high is the highest 99th percentile for the whole range, cumulative returns are the sum of log exchange rate returns. The event day is the day of the extreme risk reversal reading as a proxy for extreme speculative sentiment.  If speculation is uninformed, extremes should be followed by reversals; if speculation is informed, extremes are more likely to be followed by a random walk.}
\label{fig:ES1}
\end{figure}
% This description provides an explanation for the observation from \citet{ShillerVol} that stock prices are more volatile than the underlying earnings information on which valuation should be based. In addition, \citet{Delong1990noise} and \citet{shleifer1997limits} show that constrained horizons can cause noise-trading to limit the ability of informed traders to take advantage of profitable opportunities for fear of further divergence from fundamental value before realignment is achieved.

The nature of speculation is central to many contemporary questions about international financial markets. For example, regulations or taxes that reduce the amount of speculation may be beneficial if they limit risk-taking and encourage more long-term attitudes towards investment.  If speculation is largely uninformed it will add to the level of noise and make it even harder for informed investors to overcome uncertainty and apply fundamental analysis about long-term value; will increase the tendency towards activity that is focused on the short-term.  However, if speculation is informed and part of the process of price discovery, regulatory measures to reduce its influence may also reduce liquidity and undermine the prices of price discovery. Increased volatility, heightened risk and reduced transparency together with other costs may provide a counteraction to any positive benefits.

The rest of the paper proceeds as follows:  Section Two discusses the distinction between informed and uninformed speculators; Section Three talks about the measurement of speculation and the testing of extreme positions; Section Four reviews the evidence; Section Five concludes with a discussion of policy implications.

\section{Informed and uninformed traders}
Speculation plays an  important role in the areas of market-microstructure, the rationale for the trading of financial securities as well as providing the explanation for rational expectations are supposed to operate. However, there is no clear, unambiguous theory of speculation.  Speculation is frequently associated with the term \emph{trading}, but trading is broader as it encompasses market-making. A distinction made between the short-term orientation of speculation and trading and the more long-term activity of investment.   Speculation and trading can be categorised as being informed or uninformed. Those trading without information are also called \emph{noise-traders} or \emph{liquidity traders}.  However, the terms are not fixed and  \citet[p. 1]{GortonNoise} categorise noise-traders as all those trading in markets for ``non-information-based reasons''. 
%these two paragraphs above may go later and just a very short coverage of the distinction between noise-traders and liquidity traders. 
The distinction between informed and uninformed traders is a central component of microstructure research.  The focus of the \citet{Kyle1985Continuous} model is a market-maker receiving orders from the informed and the uninformed.  On the assumption that uniformed orders arrive in a random fashion, the weight of orders implies something about the knowledge of the informed and should led to adjustment of the bid-ask spread as part of the process of price discovery.  This bid-ask spread is the payment that the market-maker receives from the uninformed in return for taking the risk of meeting an informed trader.  In a similar way \citet{glosten1985bid}describes a model where the bid-ask spread is a function of adverse selection, being caused by the risk that orders come from informed traders.

Uninformed activity provides a rationale and the opportunity for trading. Unless there is a difference of opinion about value, there is no reason to trade. Uninformed speculation can create space between price and fundamental value that can be exploited by informed speculators.  This space can encourage informed traders to pay the cost of acquiring information.  \citet{Grossman1980Impossibility} argue that there is equilibrium between the cost of obtaining information and the benefits that are expected to come from the use of the information.

@@

\citet{EasleyPIN} use information as an explanation for the effect of large orders on stock prices.  The \emph{order effect} may be a consequence of the need to adjust inventories when a market-maker receives a large order; that effect may also be a function of the market-maker knowledge that large orders are more likely come from to informed traders seeking to realise their informational advantage.  \citet{EasleyPIN2} develop a technique to estimate the probability that trading is informed.  The find that the high liquidity stocks have the highest level of information and the most frequent arrival rate of informed traders but that price changes are limited by the volume of uniformed traders. Low liquidity stock, by contrast, have fewer information events but informed traders have a much larger effect on prices.

Informed speculation is also used as the explanation for the tendency of Rational Expectations Equilibrium (REE) to prevail in the long-run.  Those economic agents able to determine the fundamental value or discern the underlying economic model will be able to speculate and drive prices towards equilibrium.   This tends to leave uninformed speculation as a pure noise that can be characterised as white-noise random variable\footnote{See \citet{muth1961Rational} for the original adjustment to a cobweb model.}. While this construction may be justified for the analysis of microstructure and market efficiency, it is an over-simplification that is inappropriate when seeking more understanding of speculation itself.  Institutional and psychological explanations for noise-trading frequently suggest that there are endogenous non-random elements to speculative activity.
Noise-trading is associated with psychological biases and restricted information while liquidity-trading is about institutional constraints.  Therefore, noise is associated with behavioural issues, over-confidence and the use of irrational or purely fashionable investment tools. See \citet{DeBondtAnalysts} for an analysis of over-reaction and the \emph{representativeness heuristic}, \citet{OdeanOC} for a study of over-confidence and  \citet{ShillerFashion} for an assessment of investment fads.  Traders without information may not make decisions independently of others. Information may be inferred from the buying and selling conducted by other traders\footnote{See for example \citet[p.101]{Keynes1936} and the beauty contest.}.

Institutional constraints may mean that economic agents just execute customer orders at the current price or it may be that they buy and sell with reference to a benchmark index or to fulfil a hedging strategy. The institutional explanations are associated with the allocation of credit for investment and trading purposes:  as asset values decline exchanges, banks and other providers of liquidity require larger margins and deeper haircuts; investors liquidate their fund holdings; the attraction of hedging increases.  See \citet{BrunnermeierLiquidity} and \citet{Gorton} for some discussion of the way that margin-calls and collateral constraints can force investment decisions.

\citet{FriedmanPositive} said that the informed market participants would drive the uninformed from the market.  However, a large proportion of uninformed traders in a market can undermine this process.  \citet{Delong1990noise} and \citet{shleifer1997limits} have presented models where \emph{noise-trader risk} can prevent the arbitrate of informed speculators from driving prices back towards the fundamental value.  Noise-trader risk is the risk that noise-traders will drive prices even further away from equilibrium before convergence. This can limit the activity of informed speculators when they are risk averse.\footnote{This three-period, overlapping generation model with informed and uninformed traders has a price equation \begin{equation}
p_t = 1 + \frac{\mu(\rho_t - \rho^*)}{1 + r} + \frac{ \mu \rho^*}{r} - \frac{2 \gamma \mu^2 \sigma_{\rho}^2}{r(1 + r^2)}.
\label{eq:NP2}
\end{equation} 
 where $p_t$ is the price; $\mu$ is the share of uninformed traders; $\rho^*$ is the average misperception of uninformed traders; $\rho_t$ is the price mis-perception of the uninformed traders at time t; $\gamma$ is the level of risk aversion; $\sigma^2_{p,t+1}$ is the variance of the misperception.}

@@

\subsection{Empirical work}
Empirical assessment of speculation has been limited by the scarcity of data.  It is not very easy to distinguish between speculators and other types of investors.  It is even harder to distinguish between the informed and uninformed.  One approach has been to assume that noise-traders utilise simple investment strategies and to use tests of these strategies as a means of testing speculation.  Amongst the strategies that have been used in this way are momentum, contrarian and carry trades.  However, as these techniques can be profitable, it is not clear that these can be identified as the domain of the uniformed.   It is also possible to assume that certain investors usually have more information than others.  For example, hedge funds, investment banks and other professionals can be assumed to be informed while retail or day traders may be uninformed.

\citet{EasleyPIN2} propose a model to try to measure informed market activity based on the assertion that the bid-ask spread is a function of the level of informed trading.  Large orders are assumed to be informed as informed traders are most likely to want to swift use of their informational advantage.\footnote{With market-makers using Bayes' Rule to update their fundamental valuation by using order flow to estimate the joint probability that there has been an information event and that it has been good news or bad news, the authors seek to operationalise the following model:
\begin{subequations}
\begin{equation}
b(t) = E[V_i|t] - \frac{\mu P_b(t)}{\varepsilon + \mu P_b(t)}(E[V-t|-\underline{V}])
\end{equation}
\begin{equation}
a(t) = E[V_i|t] + \frac{\mu P_g(t)}{\varepsilon + \mu P_g(t)}(\overline{V}E+ [V-t|])
\end{equation}
\end{subequations}
  where $b(t)$ is the bid price at time $t$; $V$ is the value of the security; $\mu$ is the share of informed traders in the market; $P_g(t)$ is the probability that there has been some good news and $P_b(t)$ is the probability that there has been some bad news; $\varepsilon$ is the share of uninformed traders in the market; $\underline{V}$ is the value if there is bad news and $\overline{V}$ is the value if there is good news.}

However, identifying large orders and making an assessment of the level of informed trading is much harder in an over-the-counter market like foreign exchange. While electronic order books have become more widely used, much of the market-making still takes place through bilateral agreements and information about the nature of transactions including size and origin of orders is not widely available.\footnote{\citet{Lyons1995Microstructure} and \citet{Evans2002Order} have gained access to proprietary data and have used the initiation of orders as an explanatory variable in an equation to explain the evolution of major exchange rates.  They find that $\$$ 1 billion of net US dollar purchase moves the Deutsche mark price of the US unit by 0.5 percent. They argue that the orders crystallise a range of information that is used to value currencies.}  \citet{LeeandReady} seek to use price action to infer the nature of the order that has been placed into the market.  The \emph{tick-test} has four classifications: the up-tick, the down-tick, the zero-up-tick and the zero-down-tick. A trade is an up-tick if the price is higher than the previous trade.  If the price is the same as the previous trade, if the previous trade was classified as an up-tick, this trade is a zero-up-tick. Under this system, the up-ticks are assumed to be buy orders and the down-ticks are sell orders. Up-ticks are classified as buy orders and down-ticks are classified as sell orders.  These estimates of the market activity can be used to estimate the level of informed trading.  
\section{Measuring speculation and the event study}
The aim of this study is to assess whether speculation in the foreign exchange market is informed or uninformed. Speculation must be identified and then extreme speculation will then be compared with what would be expected if it were informed or uninformed.  If informed and part of the price-discovery process, extreme speculation should drive the price towards a new equilibrium; the following price performance should be a random walk.  If speculation is uninformed, extremes should cause a divergence from fundamental value and a price reversal should be anticipated.

The \citet[pp. 14-15]{Delong1990noise} model is summarised by Equation \ref{eq:NP2}. There are two parameters that can be used to measure speculation:  the price misperception $(\rho)$ and the weight of noise-traders in the market $(\mu)$.  If speculation is uninformed and either of these parameters reach an extreme, the model indicates that prices should subsequently reverse.\footnote{See \citet{Hayward2013} for full details of simulations of this model.} Therefore, if misperception or weight of speculation can be measured, this assertion can be tested.

\subsection{Measuring the intensity of speculative sentiment}
The measurement of speculative intensity must identify how far speculators believe that the market will move in a particular direction. In the De-Long model, this would mean that for Equation \ref{eq:NP2}, it is the extreme deviation of $\rho$ from one.  In the absence of a daily survey of speculators, this study uses a unique series of \emph{risk-reversal skew} as an objective measure of speculative sentiment.\footnote{It appears that this is the only period for which the data currently exist.  The risk-reversal data is very difficult to obtain in an over-the-counter market like foreign exchange where there is no centralised exchange to record and collect information, generally being the property of quoting institutions.  The risk-reversals are the average prices quoted at 10:00 GMT by major banks on the Thomson-Reuters composite option pages (OPT1)  for the period between January 2 1996 and September 6 2002.  The risk-reversals are the spread of implied volatility for a 1 month 25 delta at-the-money call option relative to the equivalent put.  Risk-reversal skew is the relative price of currency put and call options for the same strike price.}  The \emph{Put-Call-Parity} in the  \citet{Black1973Options} option pricing model assumes that these will be the same, in practice they are not\footnote{While the Black-Scholes model is the basis for valuing options, where the price of a non-dividend paying stock trades at a discount to the forward rate equal to the risk free rate, the forward rate in a currency option is determined by the relative interest rate differential between the two currencies.   The \citet{GarmanKohlhagan} adjustment that accounts for the difference in interest payments on the two currencies is used.   The spot discount is determined therefore by the relative interest rate differential for the appropriate maturity.   This model allows the market practice of pricing options from the volatility implied by the model.  In other words, the Garman-Kohlhagan is used to extract the volatility that is implied by a specific option price given the strike and other parameters.}  A call option is the right to buy and a put option is the right to sell.  The delta of an option measures the effect of a change in the spot price on the value of the option.  This can range from zero for deeply out-of-the-money options to unity for deeply in the money strikes.  The at-the-money delta is 50 as there is a 50:50 chance of the option being exercised.  The benchmark for option risk-reversals is 25 delta.  This is a measure of how far the option is out of the money or how far the strike is from the forward rate.

If the exchange rate is expected to follow a random walk with drift towards the 1 month forward, implied volatility on the put and the call for the same delta should be equal.  Any deviation from this equivalence is an indication that a deviation from the random walk with drift is implied in the market price of options.\footnote{If the risk-reversal for a 1-month USD/CAD is quoted at 0.15-0.28\% and the implied volatility is 8.50\%, the market-maker would be willing to buy the 1-month 25 delta USD-put-CAD-call at 8.65\% and sell the USD-call-CAD-put at 8.50\%.  The dealer pays 0.15\%.  Alternatively, the dealer would sell a 25 delta call at 8.78\% and buy the USD-call-CAD-put at 8.50\%, earning 0.28\% \citet{Global}. }  Therefore, the relationship between implied volatility for calls and implied volatilities for puts for the same delta and maturity is an indication of how much more expensive calls are relative to the equivalent put or as the market bias towards puts over calls and as an indication of market sentiment.  The larger the absolute level of the risk-reversal skew, the greater the relative price and, if the deviation is based on information, subsequent price action should be zero on average; if the deviation is based on noise, there should be a good chance of a reversal.
